---
title: "Project4 Group10"
output: html_notebook
---

### Step 1 Load Data and Train-test Split
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
data <- read.csv("../data/ml-latest-small/ratings.csv")
set.seed(0)
test_idx <- sample(1:nrow(data), round(nrow(data)/5, 0))
train_idx <- setdiff(1:nrow(data), test_idx)
data.train <- data[train_idx,]
data.test <- data[test_idx,]
```

### Step 2 Matrix Factorization
#### Step 2.1 load the function for A2, P2 and P3
```{r}
source("../lib/Probalistic_ Matrix_Factorization_w_cv.R")
```

### Step 2.2 pre Tunning for single A2 to shrink the possible parameters

We use pre-fine-tunning to shink the number of possible parameters for our future model. And as we fixed the sigma, only parameter that would infect the model is D, sigma_U, sigma_V and lrate.

So we choose three candidate value for D, sigma_U, sigma_V and lrate.
```{r}
# tuned parameters
t_D <- c(5, 10, 15)
t_sigma_V <- c(.1, .5, 1)
t_sigma_U <- c(.1, .5, 1)
t_lrate <- c(.0001, .0005, .001)

# Do the first step of CV(on the A2) to select parameters that can be used 
parameters <- expand.grid(D = t_D, sigma_V = t_sigma_V, sigma_U = t_sigma_U, lrate = t_lrate)
```

Do cross validation
```{r eval=FALSE}
cv_train_rmse <- matrix(NA, dim(parameters)[1], 20)
cv_test_rmse <- matrix(NA, dim(parameters)[1], 20)

for(i in 1:dim(parameters)[1]){
  tmp_D <- parameters$D[i]
  tmp_sigma_V <- parameters$sigma_V[i]
  tmp_sigma_U <- parameters$sigma_U[i]
  tmp_lrate <- parameters$lrate[i]
  
  ret <- cv_A2.function(data, data.train, K = 5,
                        D = tmp_D, sigma_V = tmp_sigma_V, sigma_U = tmp_sigma_U, lrate = tmp_lrate)
  cat("--------------------------------------- No.",i,"Done --------------------------------------------", "\n")
  cv_train_rmse[i,] <- ret$mean_train_rmse
  cv_test_rmse[i,] <- ret$mean_test_rmse
}
cv_A2_train_rmse <- cv_train_rmse
cv_A2_test_rmse <- cv_test_rmse
colnames(cv_A2_train_rmse) <- sapply(c(1:20), function(x) paste("train_",x,sep = ''))
colnames(cv_A2_test_rmse) <- sapply(c(1:20), function(x) paste("test_",x,sep = ''))

save(cv_A2_train_rmse, file = "../output/cv_A2_train.RData")
save(cv_A2_test_rmse, file = "../output/cv_A2_test.RData")
```

For each parameter, 2e select top two possible value which can make model to have the smallest mean(RMSE).
```{r}
load("../output/cv_A2_train.RData")
load("../output/cv_A2_test.RData")

cv.return <- parameters %>% cbind(cv_A2_train_rmse, cv_A2_test_rmse)

pre_ft_D <- (cv.return %>%
  group_by(D) %>% 
  summarise(test_mean = mean(test_20)) %>% 
  arrange(test_mean) %>% 
  select(D) %>% as.matrix() %>% c())[1:2]

pre_ft_sigma_U <- (cv.return %>%
  group_by(sigma_U) %>% 
  summarise(test_mean = mean(test_20)) %>% 
  arrange(test_mean) %>% 
  select(sigma_U) %>% as.matrix() %>% c())[1:2]

pre_ft_sigma_V <- (cv.return %>%
  group_by(sigma_V) %>% 
  summarise(test_mean = mean(test_20)) %>% 
  arrange(test_mean) %>% 
  select(sigma_V)%>% as.matrix() %>% c())[1:2]

pre_ft_lrate = (cv.return %>%
  group_by(lrate) %>% 
  summarise(test_mean = mean(test_20)) %>% 
  arrange(test_mean) %>% 
  select(lrate)%>% as.matrix() %>% c())[1]
```

Then we draw the graph to see the convergency of these 8 model sets.
```{r}
cv.grap <- cv.return %>% 
  filter(D %in% pre_ft_D,
         sigma_U %in% pre_ft_sigma_U,
         sigma_V %in% pre_ft_sigma_V,
         lrate == pre_ft_lrate) %>% 
  mutate(param_set = paste("D =",D,", sigma_U =", sigma_U, ", sigma_V =", sigma_V)) %>% 
  select(-c(D,sigma_V,sigma_U,lrate)) %>% 
  pivot_longer(-c(param_set), names_to = "Train_Test", values_to = "RMSE") %>% 
  separate(Train_Test, into = c("Train_Test", "Epoch"), sep = "_") %>% 
  mutate(Epoch = as.integer(Epoch))


a <- unique(cv.grap$param_set)[1:4]
b <- unique(cv.grap$param_set)[5:8]

cv.grap %>% 
  filter(param_set %in% a) %>% 
  ggplot(aes(x = Epoch, y = RMSE, col = Train_Test)) + geom_point() + facet_grid(~param_set)


cv.grap %>% 
  filter(param_set %in% b) %>% 
  ggplot(aes(x = Epoch, y = RMSE, col = Train_Test)) + geom_point() + facet_grid(~param_set)
```

```{r, echo=FALSE}
best_param_sets <- expand.grid(D = pre_ft_D, sigma_U = pre_ft_sigma_U, sigma_V = pre_ft_sigma_V)

cat("Promising value for D is ", pre_ft_D[1], " and ", pre_ft_D[2], ".", "\n", sep = "")
cat("Promising value for sigma_U is ", pre_ft_sigma_U[1], " and ", pre_ft_sigma_U[2], ".", "\n", sep = "")
cat("Promising value for sigma_V is ", pre_ft_sigma_V[1], " and ", pre_ft_sigma_V[2], ".", "\n", sep = "")
cat("We fix the learning rate to be ", pre_ft_lrate, ".", sep = "")
```

